---
title: "Coinbase price regression"
output:
  html_document:
    df_print: paged
---

The prices are Coinbase prices maintained as one of datasets here:

http://api.bitcoincharts.com/v1/csv/coinbaseUSD.csv.gz

Warning: the compressed CSV file is currently heavier than 147 MB.

```{r load_data}
tmp <- tempdir()
filename <- paste0(tmp, "/coinbaseUSD.csv.gz")
download.file("http://api.bitcoincharts.com/v1/csv/coinbaseUSD.csv.gz", filename)
coinbase <- read.csv(gzfile(filename), header = F)
colnames(coinbase) <- c("timestamp", "price", "xxx")
```

Among many models I checked, the one fitting a parabole to the logarithm of the price got me the best fit, with coefficient of determination higher than 0.97.

```{r fit_model}
model <- lm(log(price) ~ poly(timestamp, 2), coinbase)
```

We can use the fitted model to predict the future trendline values of the BTC price and times of crossing the 100k and 1M barriers.

```{r plotting}
# Future timestamps
predictionstamps <- as.double(as.POSIXlt(unlist(lapply(2014:2020, function(year)
  paste(year, 1:12, "01 00:00:00", sep = "-")))))

# Bitcoin price values for future timestamps
predictions <- exp(predict(model, data.frame(timestamp = predictionstamps)))

# Plotting
plot(coinbase$timestamp, coinbase$price, type = "l",
  xlim = c(min(coinbase$timestamp), max(predictionstamps)),
  ylim = c(1e2, max(predictions, coinbase$price)), log = "y",
  xlab = "Year", ylab = "Price (log)",
  main = "Coinbase price and a trendline for a log-parabolic model",
  axes = F)

# Axes
yearstarts <- as.double(as.POSIXlt(paste(2014:2020, "01-01 00:00:00", sep = "-")))
axis(1, at = yearstarts, labels = 2014:2020)
axis(2, at = c(1e2, 1e3, 1e4, 1e5, 1e6, 1e7), c("100", "1k", "10k", "100k", "1M", "10M"), las = 2)

abline(v = yearstarts, lty = 2, col = "lightgray")
abline(h = c(1e3, 1e4, 1e5, 1e6), lty = 2, col = "lightgray")

# Add the trendline
lines(predictionstamps, predictions, type = "l")

# Add plot creation time
text(1.45e9, 1e10, paste("Updated", Sys.Date()), pos = 4)

# Display coefficient of determination
r2 <- signif(summary(model)$r.squared[1], 2)
text(predictionstamps[length(predictionstamps)], 0.9 * predictions[length(predictions)],
     bquote(R^2 == .(r2)), pos = 2)
```

Bitcoin is growing very high above the trendline at the moment!

```{r residuals}
coinbase$residuals <- residuals(model)
```